---
title: "Learning From Data 112-2 Documentary"
author: 
  - CHIH-TSE LIU
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    fig_width: 8
    fig_height: 3.5  
    latex_engine: "xelatex"
    includes:
      in_header: "Rnote.tex"
    keep_tex: true
header-includes:
  - \renewcommand{\and}{\\}
fontsize: 12pt
abstract: 
  This.
---

\begin{center}
Table of Contents 
\label{sec:TOP}
\begin{itemize}
\item \hyperref[sec:PRE]{Packages and Functions}
\item \hyperref[sec:DAT]{Data Overview}
\item \hyperref[sec:VAR]{Variable Inspection}
\item \hyperref[sec:EDA]{Exploratory Data Analysis}
\item \hyperref[sec:]{}
\item \hyperref[sec:]{}
\item \hyperref[sec:]{}
\end{itemize}

% '\thispagestyle{empty}'  surpresses page number for the cover
% '\pagenumbering{arabic}' resets page number
\end{center}

\thispagestyle{empty} 
\clearpage
\pagenumbering{arabic}

\externaldocument{Outline}

```{r setup, include = FALSE}
# hide warnings and messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# prevent R output from exceeding page margin
# not all R functions respects this setting
options(width = 75)

# set graph size
#knitr::opts_chunk$set(out.height = "50%", out.width = "100%")

# 75 is the optimal maximum limit for characters
#########-#########-#########-#########-#########-#########-#########-#####
```

\newpage
\section{Packages and Functions Specified}
\label{sec:PRE}
\hypertarget{sec:PRE}{}
## Packages
```{r}
# for error normality test (MLR.6)
library(nortest)

# for homoskedasticity test (MLR.5)
# library(alr3)

# for box-cox transformation (MLR.5, MLR.6)
library(MASS)

# for data frame manipulation
library(dplyr)

# for EDA
library(ggplot2)

# for GLS
library(nlme)

# for model selection
# library(ALSM)
library(leaps)
library(olsrr)

# for heatmaps and graphs
library(pheatmap)
library(PerformanceAnalytics)

# for large data sets manipulation
library(data.table)

# for graphing
library(lattice)
library(RColorBrewer)
library(stringr)

# 
library(corrplot)
```
